---
title: "p8105_hw2_zs2539"
author: "Zhengwei Song"
date: "`r Sys.Date()`"
output: github_document
chunk_output_type: console
---

# Problem 1 Solution

## Importing data tables
```{r}
library(tidyverse)
NYC_transit_data_raw = read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
            col_types = cols(
    Route8 = col_character(),
    Route9 = col_character(),
    Route10 = col_character(),
    Route11 = col_character()
        ))
```

## Cleaning up variable names
```{r}
names(NYC_transit_data_raw)
NYC_transit_data_clean_names = janitor::clean_names(NYC_transit_data_raw)
names(NYC_transit_data_clean_names)
```

## Selecting required columns in problem 1
```{r}
library(dplyr)
NYC_transit_data_selected_cols = select(NYC_transit_data_clean_names, line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada)
```

## Combining routes columns
```{r}
NYC_transit_data_combined_cols = 
  pivot_longer(
    NYC_transit_data_selected_cols, 
    route1:route11,
    names_to = NULL,
    values_to = "route_number",
    values_drop_na = TRUE)
```

## Converting to Logical variables for entry and vending
```{r}
NYC_transit_data_logical_cols = mutate(NYC_transit_data_combined_cols,
    entry = recode(entry, "YES" = "TRUE", "NO" = "FALSE"),
    entry = as.logical(entry),
    vending = recode(vending, "YES" = "TRUE", "NO" = "FALSE"),
    vending = as.logical(vending),
    )
```

## Selecting distinct rows of the dataset
```{r}
NYC_transit_data_final = NYC_transit_data_logical_cols %>% distinct()
```

## In summary, this has read and cleaned the data as required, and the whole process is shown above. The resulting dataset is 1559 rows x 9 columns, including line, station name, station latitude / longitude, routes served, entry, vending, entrance type and ADA compliance.
```{r}
tail(NYC_transit_data_final)
```

## Answering Questions in problem 1

### It has 356 distinct stations in total by the coding below
```{r}
NYC_transit_data_final %>% distinct(station_name) %>% count()
```

### Similarly, 73 stations are ADA compliant
```{r}
NYC_transit_data_final %>% filter(ada == "TRUE") %>% distinct(station_name) %>% count()
```

### 60.38% station entrances / exits without vending allow entrance

```{r}
a = NYC_transit_data_final %>% filter(vending == "FALSE", entry == "FALSE") %>% select(station_name, entry) %>% distinct() %>% count()

b = NYC_transit_data_final %>% filter(vending == "FALSE") %>% select(station_name, entry) %>% distinct() %>% count()

a/b
```
